---
title: "Does FDI Drive Growth?"
subtitle: "Evidence from the World Development Indicators"
author: "Lucia Hernandez"
format:
  beamer: 
    theme: Szeged
    colortheme: seahorse
knitr: 
  opts_chunk:
    echo: false
    comment: NA
    message: FALSE
    warning: FALSE 

---
# Analysis
```{r}
library(tidyverse)
library(knitr)
df0 <- read_csv('WDICSV.csv')

df <- df0 %>%
  pivot_longer(
    cols = 5:70,
    names_to = 'year',
    values_to = 'score'
  ) %>%
  filter(`Indicator Code` == 'BX.KLT.DINV.WD.GD.ZS' |
          `Indicator Code` == 'NY.GDP.PCAP.KD.ZG')%>%
  pivot_wider(
    values_from = 'score',
    names_from = `Indicator Code`
  )%>%
  rename(
    fdi = 'BX.KLT.DINV.WD.GD.ZS',
    gdp = 'NY.GDP.PCAP.KD.ZG'
  )%>%
  mutate( year = as.integer(year))%>%
  group_by(`Country Name`, `Country Code`, year) %>%
  summarize(
    fdi = first(na.omit(fdi)),
    gdp = first(na.omit(gdp)),
    .groups = 'drop'
  )

```

##
```{r}
 
filter(df, 
  `Country Name` == 'East Asia & Pacific (excluding high income)'| 
  `Country Name` == 'Sub-Saharan Africa (excluding high income)'|
  `Country Name` == 'Latin America & Caribbean (excluding high income)' | 
  `Country Name` == 'Middle East, North Africa, Afghanistan & Pakistan (excluding high income)') %>%
    ggplot(aes(fdi,gdp))+
    geom_point(color = 'darkorchid4')+
    geom_smooth(method=lm, color = 'darkorchid4')+
    theme_minimal()+
    theme(plot.title = element_text(size = 12, face = 'bold'),
    strip.text = element_text(size = 12, face = "bold"))+
    facet_wrap(~`Country Name`, scales = 'free')+
    labs(
    title = 'Figure 1: FDI and GDP per capita in each region',
    subtitle = "FDI is positively associated with GDP growth in most regions, but the relationship is weakest in Latin America",
      x = "	
Foreign direct investment, net inflows (% of GDP)",
      y = 'GDP per capita (annual % growth)',
      caption = "Source: World Bank World Development Indicators (WDI)"
)

 
```

## 
```{r}
filter(df, 
  `Country Name` == 'Least developed countries: UN classification' | 
  `Country Name` == 'Low income' |
  `Country Name` == 'Lower middle income') %>%
  mutate(`Country Name` = recode(`Country Name`,
    'Least developed countries: UN classification' = 'Least developed')) %>%
  ggplot(aes(fdi, gdp)) +
  geom_point(color = 'cornflowerblue') +
  geom_smooth(method = lm, color = 'cornflowerblue') +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, face = 'bold'),
        strip.text = element_text(size = 12, face = "bold")) +
  facet_wrap(~`Country Name`, scales = 'free') +
  labs(
    title = 'Figure 2: FDI and GDP per capita least developed, low income, and middle income countries',
    subtitle = "Least developed countries show the clearest positive relationship between FDI and growth.",
    x = "Foreign direct investment, net inflows (% of GDP)",
    y = 'GDP per capita (annual % growth)',
    caption = "Source: World Bank World Development Indicators (WDI)"
  )
```
##

```{r}
df2 <- df %>%
  mutate(
    gdp_growth = if_else(gdp>0, "positive","negative"),
    fdi_gr = if_else(fdi > median(fdi, na.rm = T), "High FDI", "Low FDI"),
    fdi_group = fct_relevel(fdi_gr, "Low FDI", "High FDI"))

  

df2 %>%
  filter(!is.na(fdi), !is.na(gdp_growth))%>%
  group_by(gdp_growth)%>%
  summarize(
    n = n(),
    mean_fdi = mean(fdi, na.rm =T),
    median_fdi = median(fdi, na.rm =T)
  )%>%
  kable(digits = 2,
        caption = 'Table 1: Average FDI by GDP Growth Direction',
        col.names = c('Growth Direction', "N", "Mean FDI", "Median FDI"))
```  

 
## 
```{r}

df2 %>%
  filter(!is.na(fdi_group), !is.na(gdp_growth)) %>%
    group_by(fdi_group, gdp_growth) %>%
    summarize(n = n()) %>%
    ungroup()%>%
    group_by(fdi_group) %>%
    mutate(prop = n / sum(n)) %>%
    ggplot(aes(x = fdi_group, y = prop, fill = gdp_growth)) +
    geom_col() +
    scale_fill_manual(values = c("negative" = "azure2", 
                                 "positive" = "darkslategray3")) +
    scale_y_continuous(labels = scales::percent) +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = 'bold'))+
    labs(
      title = "Figure 3: Proportion of Country-Years with Positive vs. Negative GDP Growth",
      subtitle = "High-FDI country-years are more likely to have positive GDP growth than low-FDI country-years",
      x = "FDI Group",
      y = "Proportion of Country-Years",
      fill = "GDP Growth",
      caption = 'FDI Group calculated as being above of below the median'
    )



```
# Conclusion

## Conclusion
- Higher FDI is associated with higher GDP per capita across most regions, but the relationship is not uniform.
- East Asia and Sub-Saharan Africa show the clearest positive slopes, while Latin America shows little to no relationship. 
- Countries that receive high levels of FDI appear to have a negative slope, which could suggest diminishing returns at high inflow levels
- Least developed and lower-midle income countries show stronger positive associations than low-income countries
- Country-years with positive GDP growth had higher median FDI (1.84%) than country-yeas with negative GDP growth (0.9%). 
 



